### 2017-03-11 preproposal/preproposal.tex
Updated preproposal abstract

\begin{centering}
\Large
    \textbf{Time-series and Hierarchical Modeling for sc-RNA Expression Levels}  \\
    \vspace{2mm}
    \normalsize
    Skanda Koppula (\url{skoppula@mit.edu}), Karren Yang (\url{karren@mit.edu}) \\
    \vspace{2mm}
    \normalsize
    6.882 Project Pre-Proposal, \today \\
\end{centering}
\vspace{5mm}

Cancer cells implanted into mice exhibit patterned levels of RNA expression. The expression in each cancer cell varies over time, and the cells may naturally diverge in their expression during this time. Previous work has looked into an LDA-based clustering of expression vectors for particular time slices, resulting in interesting and biologically meaningful cell grouping shifts over time. \\

This project is interested in constructing the first generative model of this process. Specifically, we look to answer some of the following questions: 
\begin{enumerate}
\item Can we apply a topic model based probabalistic model for a collection of expression vectors for a particular time slice? Will a topic model based solution yield meaningful cell type assignments?
\item How can we apply time-series probabilistic models such as Chinese Restaurant to model cell types over time?
\item How can we model noise in expression level observations in our probabilistic models? Is it possible to use our model to establish estimates for missing coordinates in expression vectors (a common problem in high-throughput expression measurement machines)?
\item Is it possible to emulate expression data with a simple sampling processes (for the purposes of testing any models we create)?
\end{enumerate}
These were a few lines of inquiry we were interested in pursuing. The authors have access to an sc-RNA expression level dataset through their research lab.
\end{document}

>>>>

\begin{centering}
\Large
    \textbf{Bayesian Non-Parametric Models for Analyzing Single-Cell RNA Sequencing Data}  \\
    \vspace{2mm}
    \normalsize
    Skanda Koppula (\url{skoppula@mit.edu}), Karren Yang (\url{karren@mit.edu}) \\
    \vspace{2mm}
    \normalsize
    6.882 Project Pre-Proposal, \today \\
\end{centering}
\vspace{5mm}

Cancer cells injected in mice grow into tumors composed of distinct sub-populations of cells, which exhibit patterned levels of gene expression. In order to characterize sub-populations of tumor cells and identify useful gene modules, we propose applying Bayesian nonparametrics to model single-cell RNA-sequencing (scRNA-seq) data from mouse tumors. Parametric topic models based on matrix factorization have shown promise in identifying meaningful clusters of cells and genes in heterogeneous cell populations. However, the need to pre-determine the number of clusters is a major limitation, particularly when extending these models to time-series data, as sub-populations of cells may appear, disappear, or diverge in expression over time. We propose exploring the utility of nonparametric models based on the Hierarchical Dirichlet Process (HDP) and the Indian Buffet Process (IDP) for modeling scRNA-seq data from single time points. We plan to compare their performance to that of existing methods for scRNA-seq analysis, as well as that of the top parametric topic model, Latent Dirichlet Allocation (LDA). Time permitting, we would also like to explore the extension of these models to time-series data, in order to understand how sub-populations of cells in the tumor microenvironment evolve over time.

\end{document}
\grid
\grid
